An Overview of the Current State of Test-First vs. Test-Last Debate

### 2013-11-24 Paper_Draft.tex
Finished rewriting data section and first paragraph of discussion.

\subsection{Waterfall Development and Testing}
\mycomment{present tense this paragraph}
A popular software development model that was developed in the 1970's is the waterfall software development model, where software is developed in a series of phases.  The waterfall development model is popular because it is simple to implement correctly and is time efficient. The first phase in the standard waterfall model is the requirements phase, where requirements are set by the customer or design company.  Next the design phase occurs in which the product is designed. The product is then built in the implementation phase.  The final phase is the verification phase where testing and debugging occur ~\cite{wiki:xxx}.  The phases were set up in this manner to have previous phases make later phases simpler to complete.  Due to the fact that these processes often put testing at the end of development, a certain type of testing method, test-last testing, was the only testing method that made sense to use.
http://start.fedoraproject.org/
Test-last testing is currently a popular testing method and is usually the first testing method that people tend to implement.  Test-last testing is the practice of writing tests after code has been written to check the functionality of the written code.  These tests are then used by the developer to fix their code until no further errors are found by the tests.

@@

\subsubsection{Review by Jeffries and Mitnik}

In 2012, Hellmann et. al conducted a summary on the current status of research of TDD.  Within Hellmann's background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.  Additionally while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.

@@

@@

\begin{itemize}
\item which approach they would choose in the future (\textbf{Choice})
\item which approach was the best for the project(s) they
completed (\textbf{BestApproach})
\item which approach would cause them to more thoroughly
test a program (\textbf{ThoroughTesting})
\item which approach produces a correct solution in less
time (\textbf{Correct})
\item which approach produces code that is simpler, more
reusable, and more maintainable (\textbf{Simpler})
\item which approach produces code with fewer defects

@@

The results of the study are shown in figure 1.  The results were divided between a beginner group (students within their first few code classes) and a mature group (students nearing graduation and professionals).  In the results section Janzen concluded that beginner coding students have a preference toward test-last testing while more mature coding students and industry professionals show a preference toward test-first testing.

\subsection{Analysis}
Two of the three studies covered the topic of code coverage.  In both studies by Kollanus and Lemos et al. it was acknowledged that increased test coverage had occurred in test-first testing methods compared to their test-last counterparts.  All three papers brought up total development time and noted that test-first testing tended to increase development time, although both Kollanus's review and Hellmann's review provide notable contradictory studies to this conclusion.   All three studies in one way or another also brought up code correctness in the form of external code quality in Kollanus or in the form of code quality in the case of Hellmann et al.  In this case the results were across the board where Lemos et al. stated that there was no code correctness difference between test first and test last whereas Kollanus said there was weak evidence that test-first testing, in the form of TDD, produces more correct code then test-last testing.  Hellmann et al provided contradictory evidence backing that in most studies test-first testing has shown better code correctness then test last testing but there was a noted study where test-first testing was less code correct then test-last testing.   In summary, with the studies given, test-first testing produces more code coverage and is likely to take more time to code then test-last testing but due to contradictions in the research given we can not draw a conclusion on code correctness.

\subsection{Discussion}
Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010} , including the literary review done by Kollanus ~\cite{Kollanus:2010}, have noted that it is hard to draw conclusions with the given research because that many studies that compare test-first to test-last testing are contradictory to one another.  For example in ~\cite{Hellman:2012},  there is a mention of previous summary studies that have had issues with conflicting data.  This is then followed by a series of examples of data that has conflicted in other studies in the past.  In Kollanus's literature review ~\cite{Kollanus:2010}, Kollanus points out that many articles in her study contradicted other articles which made it hard to make sound conclusions from the research given and may be a confounding factor in her study.  This seems to occur because of two main factors: participant experience and improper study implementation and documentation.

In the case of participant experience there seems to be a skill gap between test first and test last testing.  In an opinion study by ~\cite{Janzen:2007} it was found that a majority college students said they were not comfortable with and had trouble with comprehending TDD, even after doing a coding exercise with it.  The majority of students also said that they were much more comfortable implementing test last methods and would prefer to use test last methods in the future.  Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.  They also stated they felt it was relatively easy to understand while students did not. In another opinion study ~\cite{Kettunen:2010}, by Kettunen et al, a common theme throughout the paper was examples of individual testimonies and research conclusions that pointed to test first testing being harder to implement than test last testing.   Considering that most conducted experiments focus on college students and industry professionals, who have vast experience differences, the fact that test first testing may be harder to implement for college students compared to industry professionals may create differing results between studies.   This idea is considered in ~\cite{Hammond:2012} and ~\cite{Kettunen:2010}, but no summary papers where found considering this result and comparing student participant experiment results to professional participant experiment results.

>>>>

\subsection{Waterfall Development and Testing}

A popular software development model that was developed in the 1970's is the waterfall software development model, where software is developed in a series of phases.  The waterfall development model is popular because it is simple to implement correctly and is time efficient. The first phase in the standard waterfall model is the requirements phase, where requirements are set by the customer or design company.  Next the design phase occurs in which the product is designed. The product is then built in the implementation phase.  The final phase is the verification phase where testing and debugging occur ~\cite{wiki:xxx}.  The phases were set up in this manner to have previous phases make later phases simpler to complete.  Due to the fact that these processes often put testing at the end of development, a certain type of testing method, test-last testing, was the only testing method that made sense to use.

Test-last testing is currently a popular testing method and is usually the first testing method that people tend to implement.  Test-last testing is the practice of writing tests after code has been written to check the functionality of the written code.  These tests are then used by the developer to fix their code until no further errors are found by the tests.

@@

%\subsubsection{Review by Jeffries and Mitnik}

%In 2012, Hellmann et. al conducted a summary on the current status of research of TDD.  Within Hellmann's background there is a mention of a summary study produced by Jeffries and Mitnik that found in general that TDD largely resulted in an increase in quality, but one study they identified showed instead that TDD resulted in a strong negative impact on quality.  Additionally while they showed that TDD could reduce the amount of effort required by up to 27\%, most studies found an increase in effort of up to 100\%.

@@

\begin{figure}
\centering
\psfig{file=janzengraph1.pdf,width =3in}
\caption{Results for beginner programmers in Janzen's study}
\end{figure}

\begin{figure}
\centering
\psfig{file=janzengraph2.pdf,width =3in}
\caption{Results for Mature programmers in Janzens study}
\end{figure}

@@

\begin{itemize}
\item which approach they would choose in the future (\textbf{Choice})
\item which approach was the best for the project(s) they
completed (\textbf{BestApproach})
\item which approach would cause them to more thoroughly
test a program (\textbf{ThoroughTesting})
\item which approach produces a correct solution in less
time (\textbf{Correctness})
\item which approach produces code that is simpler, more
reusable, and more maintainable (\textbf{Simpler})
\item which approach produces code with fewer defects

@@

The results were divided between a beginner group (students within their first few code classes) and a mature group (students nearing undergraduate graduation and professionals).  The results for the beginner group can be seen in figure 1 and the results for the mature group can be seen in figure 2.  Each graph shows the distribution of all of the participants answers to the six questions where each question is referenced by its bold keyword.  With the data given in the figures, Janzen was able to conclude that beginner coding students have a preference toward test-last testing while more mature coding students and industry professionals show no significant preference toward either testing method.

%\subsection{Analysis}
%Two of the three studies covered the topic of code coverage.  In both studies by Kollanus and Lemos et al. it was acknowledged that increased test coverage had occurred in test-first testing methods compared to their test-last counterparts.  All three papers brought up total development time and noted that test-first testing tended to increase development time, although both Kollanus's review and Hellmann's review provide notable contradictory studies to this conclusion.   All three studies in one way or another also brought up code correctness in the form of external code quality in Kollanus or in the form of code quality in the case of Hellmann et al.  In this case the results were across the board where Lemos et al. stated that there was no code correctness difference between test first and test last whereas Kollanus said there was weak evidence that test-first testing, in the form of TDD, produces more correct code then test-last testing.  Hellmann et al provided contradictory evidence backing that in most studies test-first testing has shown better code correctness then test last testing but there was a noted study where test-first testing was less code correct then test-last testing.   In summary, with the studies given, test-first testing produces more code coverage and is likely to take more time to code then test-last testing but due to contradictions in the research given we can not draw a conclusion on code correctness.

\subsection{Understanding Contradictions}
Many summaries of TDD research ~\cite{Hammond:2012, Hellman:2012, Kettunen:2010, Kollanus:2010} have noted that studies that compare test-first to test-last testing tend to contradict one another.  Kollanus points out that articles in her review contradicted other articles which made it hard to make sound conclusions from the research given, especially on the subject of productivity.  The amount of contradictions concerned Kollanus enough that she noted the amount of contradictions as a potential confounding factor for her results.  Contradictions also existed between our review by Kollanus and our experiment by Lemos et al.  Although both articles seem to agree that code coverage is increased with test-first methods neither study seems to totally agree on the attributes of code correctness or total development time.  Since contradictions occur often in the results of our study we considered it to be important to explore some current theories on why these contradictions exist.  In particular we will be exploring two potential causes: participant experience and improper study implementation/ documentation.

In the case of participant experience there seems to be a skill gap between test first and test last testing.  This gap is best shown by the results from Janzen  Coding professionals, on the other hand, actually preferred test first programming and were much more comfortable with it than students.  They also stated they felt it was relatively easy to understand while students did not. In another opinion study ~\cite{Kettunen:2010}, by Kettunen et al, a common theme throughout the paper was examples of individual testimonies and research conclusions that pointed to test first testing being harder to implement than test last testing.   Considering that most conducted experiments focus on college students and industry professionals, who have vast experience differences, the fact that test first testing may be harder to implement for college students compared to industry professionals may create differing results between studies.   This idea is considered in ~\cite{Hammond:2012} and ~\cite{Kettunen:2010}, but no summary papers where found considering this result and comparing student participant experiment results to professional participant experiment results.
